\chapter{Acknowledgements}
There are a great deal of people I have to thank for being able to finish this thesis. I would first like to thank my supervisor Mikael and my co-supervisor Wojciech for taking me on as a Master student even when I've had a full time job. Your patience with panicked e-mails and on the spur of the moment requests for zoomm meetings is deeply appreciated. You have both taught me a lot, and I look back fondly on the summer I was lucky enough to join you both in Japan for the 4D space Numerical Workshop.

I would like to thank all the great scientists and programmers who contributed to PINC and made it easier to navigate a complex code base with helpful comments; Vigdis Holta, Gullik Killie, Sigvard Marholm and Jan Deca thank you for creating an excellent piece of simulation software and for helping me to contribute to it. I would like to extend my deepest thanks and appreciation to Steffen Brask, also a contributor to PINC, who went above and beyond anything that could be expected in helping me tackle difficult programming conundrums and in giving me ideas of how I could implement my own code in PINC. You have made me a better programmer.

Without my family, I would not have been able to suffer through a job on top being a Master student, I thank you for your endless support and for allowing me to go on long nerdy rants about physics and the Dune books. 

Finally, to my girlfriend Hélène. You kept me sane through what has been one of the toughest years of my life. Thank you so much for encouraging me, and for believing in me. 